---
title: "ARIMA"
author: "Brian Cochran"
date: "6/23/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(lubridate)
library(forecast)
```

``` {r load-data, include=FALSE}
data <- read_csv("Data/combined_data_clean2.csv") %>%
  mutate(Week=week(Date)) %>%
  mutate(Year=year(Date)) %>%
  filter(Year==2021) %>%
  filter(Watershed=="Bear Creek") %>%
  group_by(Week) %>%
  summarize(NO3_N=mean(NO3_N))

rain <- read_csv("Data/watershed_rain_data.csv") %>%
  mutate(Week=week(Date)) %>%
  mutate(Year=year(Date)) %>%
  filter(Year==2021) %>%
  group_by(Week) %>%
  summarize(Rain=mean(`Bear Creek`)) %>%
  filter(Week > 17)

data <- data$NO3_N
rain <- rain$Rain
```

<br></br>

### Observed Data

This plot shows NO3_N observations aggregated by week over the summer of 2021 in the Bear Creek Watershed.

``` {r observed-data-plot}
plot(data, xlim=c(1, 15), ylim=c(0, 7), type="l", xlab="Week", ylab="NO3_N")
```

<br></br>

### ARIMA Model

These plots show two week predictions for NO3_N based on past NO3_N and Precipitation.

``` {r arima-model}
par(mfrow=c(2,3))
for (i in 8:13){
  fit <- Arima(y=data[1:i], xreg=rain[1:i], order=c(1,1,0), seasonal=F)
  plot(data[1:i], xlim=c(1, 15), ylim=c(0, 7), xlab="Week", ylab="NO3_N", type="l")
  pred <- c(data[i], predict(fit, n.ahead=1, newxreg=rain[i+1])$pred)
  lines(x=i:(i+1), y=pred, col="red")
  lines(x=i:(i+1), y=c(data[i], data[i+1]), col="blue")
  abline(v=i, lty=3)
}
```

<br></br>

### ARIMA Model without Precipitation

These plots show two week predictions for NO3_N based on past NO3_N.

``` {r arima-model-2}
par(mfrow=c(2,3))
for (i in 8:13){
  fit <- Arima(y=data[1:i], order=c(1,1,0), seasonal=F)
  plot(data[1:i], xlim=c(1, 15), ylim=c(0, 7), xlab="Week", ylab="NO3_N", type="l")
  pred <- c(data[i], predict(fit, n.ahead=1)$pred)
  lines(x=i:(i+1), y=pred, col="red")
  lines(x=i:(i+1), y=c(data[i], data[i+1]), col="blue")
  abline(v=i, lty=3)
}
```

<br></br>

### Precipitation

This plot shows the precipitation data.

``` {r precipitation}
plot(rain[1:15], type="l", xlab="Week", ylab="Precipitation")
```




