---
title: "Watershed Project"
subtitle: ""
author: "Brian Cochran"
institute: "Cornell College"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(knitr)
library(xaringanthemer)
library(leaflet)
library(raster)
library(broom)
library(jcolors)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
watershed_data <- read_csv("../Data/combined_data_clean2.csv")
rainfall_data <- read_csv("../Data/CR_airport_rainfall.csv")
watershed_rain_data <- read_csv("../Data/watershed_rain_data.csv")
watershed_shp <- shapefile("../Data/watershed_geo/watersheds.shp")
merged_watershed_shp <- shapefile("../Data/watershed_geo/merged_watersheds.shp")
sites <- shapefile("../Data/watershed_geo/sites.shp")
```



# About the Data

- We are working with Coe College
<br></br>
- They data on Iowa watersheds
<br></br>
- This data has been collected each summer since 2002
<br></br>

``` {r data-preview, echo=FALSE}
knitr::kable(tail(watershed_data), format="html")
```
---



class: center, inverse
``` {r map, fig.height=3.5, fig.width=3.5, fig.align = 'center', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE}
# Getting max, min, and center long and lat for use in setting bounds of interactive map
temp_data <- tidy(merged_watershed_shp)
min_lat <- min(temp_data$lat)
max_lat <- max(temp_data$lat)
min_long <- min(temp_data$long)
max_long <- max(temp_data$long)
center_lat <- (min_lat+max_lat)/2
center_long <- (min_long+max_long)/2
  
# Icon for sampling site markers
siteIcon <- makeIcon(
  iconUrl="https://resources.finalsite.net/images/f_auto,q_auto,t_image_size_2/v1620195943/brentwoodk12nyus/wivrch9gpckn4nvyconr/experiment-1295041_1280.png",
  iconWidth = 25, iconHeight = 24
)

# Color palette
palette <- colorFactor("Set1", merged_watershed_shp$Watershed)
    
# Leaflet map
leaflet(width="100%", options = leafletOptions(zoomSnap = 0.25, 
                                                         zoomDelta = 0.25)) %>%
  # View and bounds
  setView(center_long, center_lat, 9.5) %>%
  setMaxBounds(min_long-0.25, min_lat-0.25, max_long+0.25, max_lat+0.25) %>%
      
  # Using ESRI NatGeoWorldMap for background
  addProviderTiles(providers$Esri.NatGeoWorldMap, options=providerTileOptions(minZoom=8.5)) %>%
      
   # Adding lines for subwatersheds
  addPolygons(data=watershed_shp, color="black", fillColor="white", opacity=0.5, fillOpacity=0, weight=1, smoothFactor = 0.5) %>%
      
  # Adding polgyons for watersheds
  addPolygons(data=merged_watershed_shp, color = "#333333", weight = 1.5, smoothFactor = 1,
              opacity = 1.0, fillOpacity = 0.5,
              fillColor = ~palette(Watershed),
              highlightOptions = highlightOptions(color = "white", weight = 2,
                                                      bringToFront = TRUE),
              label=paste0(merged_watershed_shp$Watershed, " Watershed"),
              layerId=~Watershed) %>%
      
  # Add markers and set visibility according to zoom level
  addMarkers(data=sites, label=sites$Site, icon=siteIcon, group="markers") %>%
  groupOptions("markers", zoomLevels=seq(9.5, 20, 0.25))
```
---



# Project Goals
<br></br>
- Create an interactive online dashboard to display this data
<br></br>
- Good visualizations and descriptions
<br></br>
- For a non-science/statistics audience
---



# Water Quality Index

- Five factors:
<br></br>
  - Dissolved Oxygen
  <br></br>
  - E. coli
  <br></br>
  - Total Nitrogen
  <br></br>
  - Total Phosphorus
  <br></br>
  - Turbidity
  <br></br>
  <br></br>
  <br></br>

This is based on a WQI proposed by Chris Jones, an IIHR Research Engineer
---



class: center, inverse, middle
# Dashboard
<a href="https://bcochran23.shinyapps.io/watershed_app/?_ga=2.37620064.174844941.1654812369-310770441.1654553990">Dashboard Link</a>
---



# Next Steps
<br></br>
- Continue creating visualizations
<br></br>
- Predictive models
<br></br>
- Water quality page
---



class: inverse, center, middle
# Questions?
